An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-20 paper/1.introduction.tex
modified based on the discussion with Bach-san

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
\textcolor{red}{Here, complements are minimal parts that are needed to evaluate $put$. 
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead-2},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simpley replaced by $[100,2]$ and not used to coustruct the result. In this case we can use $[..., [3]]$ as a complement.}
The key idea is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

>>>>

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
\textcolor{red}{Here, complements are minimal parts of sources that are needed to evaluate $put$. 
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simply replaced by $[100,2]$ and not used to coustruct the result. In this case we can use $[..., [3]]$ as a complement.}
The key idea is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead \circ phead \circ phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

### 2020-02-20 paper/4.xpg.tex
modified based on the discussion with Bach-san

\textcolor{red}{
In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$. At the same time, $lks$ and $lkv$ are updated with the identity functions.
This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
Thanks to this accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as possible.}
%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}

\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evlauated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$. and $\kpginline{bx_2}$}
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
Second is in $\kpginline{Case}$. We need to evaluate $lks \ s'$ and $lkv \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.

>>>>

\textcolor{red}{
In $kpg$, basically, functions for the updates will be kept in $lks$ and $lkv$.
In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by application of $lks \ s$ and $lkv \ v$. At the same time, the third argument and fourth argument are updated with the identity functions.
This evaluation is needed, because these definitions require the actual values, $es$ and $ev$.
Thanks to this accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as possible.}
%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}

\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evlauated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$. and $\kpginline{bx_2}$}
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
Second is in $\kpginline{Case}$ and not shown in the definition. We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.
